---
title: "MetaboSPAN-vignette"
author: "Andrew Patt"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{MetaboSPAN-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

MetaboSPAN is being run on a pathway network based on RaMP2. Relevant
stats about the analysis are listed in the table below.


```{r, eval=FALSE, echo = FALSE}
library(devtools)
load_all("~/RaMP-DB")
load_all("~/metabospan")
```

```{r, r setup}
library(metabospan)
library(RaMP)
library(magrittr)
library(dplyr)

pkg.globals <- setConnectionToRaMP(dbname="ramp2",username="root",conpass="mysql123",
                                   host = "localhost",socket = paste0("/lscratch/",
                       Sys.getenv("SLURM_JOB_ID"),
                       "/mysql/mysql.sock"))

## Parameters for MetaboSPAN
metabs_of_interest_cutoff=5
background_cutoff=5

load("../data/example_metabolites.rda")

```

```{r, MetaboSPAN analysis}
start <- Sys.time()
metabolite_distances <- measure_metabolite_distance(metabolites## ,
                                                    ## biospecimen = "Adipose"
)
print(Sys.time() - start)

visualize_metabolite_distances(metabolite_distances)

proximal_metabolites <- extract_proximal_metabolites(
  metabolites,
  metabolite_distances,
  ## biospecimen = "Adipose",
  absolute_cutoff = metabs_of_interest_cutoff
)

start <- Sys.time()
visualize_metabolite_network(metabolites,
  proximal_metabolites,
  ## biospecimen = "Adipose",
  label = TRUE
  )
print(Sys.time() - start)

metabolite_panel <- c(background_ids, proximal_metabolites)
metabolite_panel <- metabolite_panel[which(!is.na(metabolite_panel))]

metabolite_distances_bg <- measure_metabolite_distance(
  metabolite_panel## ,
  ## biospecimen = "Adipose"
)

proximal_metabolites_bg <- extract_proximal_metabolites(
  metabolite_panel,
  metabolite_distances_bg,
  ## biospecimen = "Adipose"
  absolute_cutoff = metabs_of_interest_cutoff
)


expanded_metabolites_of_interest <- c(
  metabolites,
  as.vector(proximal_metabolites)
)


expanded_background <- c(
  metabolite_panel,
  proximal_metabolites_bg
)


pathways_FDR <- runCombinedFisherTest(expanded_metabolites_of_interest,
  expanded_background,
  MCall = F
  )

filtered.fisher.results <- FilterFishersResults(pathways_FDR, p_fdradj_cutoff = 0.1)

plot1<-pathwayResultsPlot(filtered.fisher.results, text_size = 14, perc_analyte_overlap = 0.2, 
    min_pathway_tocluster = 2, perc_pathway_overlap = 0.2)

```

```{r, Standard analysis}
pathwaydf_default<-
    getPathwayFromAnalyte(
        analytes =
            c(metabolites),
        NameOrIds="ids",
        conpass=conpass_bw,socket=socket_bw, dbname=dbname_bw)

backgrounddf_default <-getPathwayFromAnalyte(
    analytes =
        c(as.vector(background_ids)),
    NameOrIds="ids",
        conpass=conpass_bw,socket=socket_bw, dbname=dbname_bw)

pathways_FDR_default<-
    runCombinedFisherTest(metabolites,
                          background_ids)

filtered.fisher.results <- FilterFishersResults(pathways_FDR_default, p_fdradj_cutoff = 0.1)

plot2<-pathwayResultsPlot(filtered.fisher.results, text_size = 14, perc_analyte_overlap = 0.2, 
    min_pathway_tocluster = 2, perc_pathway_overlap = 0.2)
```
# Pathway Results
## Pathways Summary table
```{r,echo = FALSE}

jaccard<-function(x,y){
return(length(
    intersect(x,
              y))/
    min(length(x),
        length(y)))
}

sig_MS_05<-pathways_FDR$fishresults%>%dplyr::filter(Pval<0.05)%>%dplyr::select(pathwayRampId)
sig_05<-pathways_FDR_default$fishresults%>%dplyr::filter(Pval<0.05)%>%dplyr::select(pathwayRampId)


stat=c("Number of significant metabolites",
       "Number of panel metabolites",
       "Similarity cutoff for significant list addition",
       "Similarity cutoff for background list addition",
       "Number of metabolites added to significant list",
       "Number of metabolites added to panel list",
       "Number of MetaboSPAN pathways with FDR pval < 0.05",
       "Number of conventional (unexpanded) pathways with FDR pval < 0.05",
       "Number of MetaboSPAN pathways with nominal pval < 0.05",
       "Number of conventional (unexpanded) pathways with nominal pval < 0.05",
       "Jaccard overlap of pathways < 0.05 between methods",
       "Time it took to generate this report (minutes)"
       )

value=c(length(metabolites),
        length(background_ids),        
        metabs_of_interest_cutoff,
        background_cutoff,
        length(proximal_metabolites),
        length(proximal_metabolites_bg),        
        pathways_FDR$fishresults%>%dplyr::filter(Pval_FDR<0.05)%>%nrow(),
        pathways_FDR_default$fishresults%>%dplyr::filter(Pval_FDR<0.05)%>%nrow(),
        pathways_FDR$fishresults%>%dplyr::filter(Pval<0.05)%>%nrow(),
        pathways_FDR_default$fishresults%>%dplyr::filter(Pval<0.05)%>%nrow(),
        jaccard(sig_MS_05,sig_05),
        difftime(Sys.time(),start_time,units="mins")
        )
summary_table<-data.frame(stat,value)

knitr::kable(summary_table,digits=2)

```

## Significant MetaboSPAN pathways (p < 0.05)
```{r}
pathways_FDR$fishresults%>%dplyr::filter(Pval<0.05)%>%
    dplyr::select(pathwayName, Pval_FDR, Pval) %>%
    dplyr::arrange(pathwayName) %>%
    knitr::kable()
```

## Significant default pathways (p < 0.05)
```{r}
pathways_FDR_default$fishresults%>%dplyr::filter(Pval<0.05) %>%
    dplyr::select(pathwayName, Pval_FDR, Pval) %>%
    dplyr::arrange(pathwayName) %>%
    knitr::kable()
```

## Metabolites added to set of interest
```{r}
if(length(proximal_metabolites)!=0){
    knitr::kable(metabospan:::source_to_common(proximal_metabolites,
                                              conpass=conpass_bw,
                                              socket=socket_bw, dbname=dbname_bw))
}

```

## Metabolites added to background
```{r}
if(length(proximal_metabolites_bg)!=0){
    knitr::kable(metabospan:::source_to_common(proximal_metabolites_bg,
                                              conpass=conpass_bw,
                                              socket=socket_bw, dbname=dbname_bw))
}
```



# Chemical 
